---
title: "GSE15573"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    code_folding: hide
    number_sections: true
    theme: lumen
---

# Background

**Title:** Transcriptome analysis describing new immunity and defense genes in peripheral blood mononuclear cells of rheumatoid arthritis patients.

El objetivo de este estudio consistió en identificar perfiles de expresión génica de pacientes con AR, con el fin de conocer sus mecanismos moleculares.

```{r libraries, include=FALSE}
library(Biobase)
library(knitr)
library(kableExtra)
library(stringr)
library(plotly)
library(ggdendro)
source("../Funciones/plotTreeEU.R")
source("../Funciones/pcaGenesEU.r")
```

# Análisis exploratorio {.tabset}

```{r load_data, include=FALSE}
load("../../Data/GSE_AR_final_selection.RData")
palette = c("#67a9cf", "#ef8a62")
# palette4 = c("#d7191c", "#fdae61", "#abd9e9", "#2c7bb6")
palette4 = c("#abd9e9", "#2c7bb6", "#fcc6c2", "#f53629")
```

**Plataforma**: Illumina human-6 v2.0 expression beadchip

**Tissue**: Peripheral blood

```{r GSE15573_muestras, echo=FALSE}
muestras <- pData(GSE$GSE15573)
expresion <- exprs(GSE$GSE15573)
muestras <- muestras[,c("status:ch1", "gender:ch1", "age:ch1")]
muestras$grupo <- apply(muestras[, c("status:ch1", "gender:ch1")], 1, function(x) {
  if (str_detect(x[1], ".*Control.*")){
    ifelse (str_detect(x[2], ".*Female.*"), "Control F", "Control M")
  } else {
    ifelse (str_detect(x[2], ".*Female.*"), "RA F", "RA M")
  }
})

tabla <- table(muestras$grupo)
kable(tabla, col.names = c("Grupo", "Muestras"))
```

 

## Boxplot

```{r GSE15573_boxplot, echo=FALSE}
Muestras <- rep(colnames(expresion), each = dim(expresion)[1])
Valores <- c()
for (columna in colnames(expresion)) {
  Valores <- c(Valores, expresion[,columna])
}
condicion <- rep(muestras$grupo, each = dim(expresion)[1])
box_data <- data.frame(Muestras, Valores, condicion)

fig <- plot_ly(box_data, y=~Valores, x =~Muestras, color=~condicion, type="box", colors=palette4) %>%
  layout(title="GSE15573 boxplot")

fig
```

## Clustering

```{r GSE15573_clustering, echo=FALSE}
correlacion <- cor(expresion)
distancia <- as.dist((1-correlacion) / 2)
cluster <- hclust(distancia)

cluster$clase <- muestras$grupo

fig <- plotTreeClust(cluster, title="GSE15573 clustering", palette = palette4)

fig
```

## PCA

```{r GSE15573_pca, echo=FALSE}
mi.pca <- pcaGenes(expresion)

mi.pca.df <- as.data.frame(mi.pca$scores)
mi.pca.df$grupo <- muestras$grupo
mi.pca.df$var.exp <- round(mi.pca$var.exp * 100, 2)
rownames(mi.pca.df) <- rownames(muestras)

fig <- plot_ly(mi.pca.df,x=~V1,y=~V2, z=~V3, text=rownames(mi.pca.df),
               color = ~grupo, colors=palette4)
fig <- fig %>% add_markers(marker= list(size=3) )
fig <- fig %>% layout(title="PCA, normalized data",
                      scene = list(
                        xaxis=list(title=paste0("PC1: ",mi.pca.df$var.exp[1],"% explained variance")),
                        yaxis=list(title=paste0("PC2: ",mi.pca.df$var.exp[2],"% explained variance")),
                        zaxis=list(title=paste0("PC3: ",mi.pca.df$var.exp[3],"% explained variance"))))
fig
```
